---
title: "Project proposal"
author: "ADA"
output: github_document
---

```{r load-packages, message = FALSE}
#install packages 
install.packages(c("janitor", "skimr"))
install.packages(c("tidyverse", "broom"))
library(tidyverse)
library(broom)
<<<<<<< HEAD
library(skimr)
## Add any additional packages you are using here

=======

## Add any additional packages you are using here
library(readr)
social <- read_csv("../data/which_social_media_platforms_are_most_popular_data_2024-11-13.csv", 
    skip = 2, n_max = 17)
# View(social)

library(readr)
demographic <- read_csv("../data/social_media_usage_among_demographic_groups - Sheet1 (1).csv", 
    skip = 1)
# View(demographic)
>>>>>>> 1444b959a3f7cb906972d22331bbbf5abe1da545
```

## 1. Introduction

The introduction should introduce your general research question(s) and your data (where it came from, how it was collected, what are the cases, what are the variables, etc.).

Text goes here. Refer to the Markdown Quick Reference: Help -> Markdown Quick Reference. 

## 2. Data

Text goes here. Place your data in the /data folder, and add dimensions and codebook to the README in that folder. Then print out the output of glimpse() or skim() of your data frame.

```{r}
# Code goes here
# Read in your data file
setwd("/cloud/project/proposal")
library(readr)
social <- read_csv("../data/which_social_media_platforms_are_most_popular_data_2024-11-13.csv", skip = 2, n_max = 17)

demographic <- read_csv("../data/social_media_usage_among_demographic_groups - Sheet1 (1).csv", skip = 1)
# Print the output of glimpse() or skim()
demographic <- read_csv("../data/social_media_usage_among_demographic_groups - Sheet1 (1).csv", skip = 1)

<<<<<<< HEAD
glimpse(demographic)
skim(demographic)
=======
>>>>>>> 1444b959a3f7cb906972d22331bbbf5abe1da545

```

## 3. Data analysis plan

Text goes here.
- What variables will you visualize to explore your research questions?
- Will there be any other data that you need to find to help with your research question?
- Very preliminary exploratory data analysis, including some summary statistics and visualizations, along with some explanation on how they help you learn more about your data. (You can add to these later as you work on your project.)
- The data visualization(s) that you believe will be useful in exploring your question(s). (You can update these later as you work on your project.)


```{r}
# Code goes here
# Code to calculate summary statistics
# Code for a visualization
```
