---
title: "Project proposal"
author: "ADA"
output: github_document
---

```{r load-packages, message = FALSE}
#install packages 
install.packages(c("janitor", "skimr"))
install.packages(c("tidyverse", "broom"))
library(tidyverse)
library(broom)
library(skimr)
## Add any additional packages you are using here

## Add any additional packages you are using here
library(readr)
social <- read_csv("../data/which_social_media_platforms_are_most_popular_data_2024-11-13.csv", 
    skip = 2, n_max = 17)
# View(social)

library(readr)
demographic <- read_csv("../data/social_media_usage_among_demographic_groups - Sheet1 (1).csv", 
    skip = 1)
# View(demographic)
```

## 1. Introduction

Our project explores trends in the popularity of various social media platforms among U.S. adults over time. The research question guiding this analysis is: *How has social media usage evolved from 2012-2024, and what patterns can be sen acros different platforms?* By examining changes in user engagement across platforms like YouTube, Facebook, Instagram, TikTok, Reddit, and others, this project aims to understand how digital behaviors and communication preferences have shifted over the past decade.

The data set used for this project was obtained from ____. The data was collected through **nationally representative surveys** of U.S. adults, asking what social media platforms they use. Each row represents survey results for a specific *year*, while the columns correspond to the **percentage of U.S. adult who reported using each platform.** This dataset provides us with a clear basis for a quantitative analysis of social media trends, allowing for comparisons across years and platforms to identify long-term evolution in digitial media engagement.


## 2. Data

Text goes here. Place your data in the /data folder, and add dimensions and codebook to the README in that folder. Then print out the output of glimpse() or skim() of your data frame.

```{r}
# Code goes here
# Read in your data file
setwd("/cloud/project/proposal")
library(readr)
social <- read_csv("../data/which_social_media_platforms_are_most_popular_data_2024-11-13.csv", skip = 2, n_max = 17)

demographic <- read_csv("../data/social_media_usage_among_demographic_groups - Sheet1 (1).csv", skip = 1)
# Print the output of glimpse() or skim()
demographic <- read_csv("../data/social_media_usage_among_demographic_groups - Sheet1 (1).csv", skip = 1)

glimpse(demographic)
skim(demographic)


```

## 3. Data analysis plan

Text goes here.
- What variables will you visualize to explore your research questions?
- Will there be any other data that you need to find to help with your research question?
- Very preliminary exploratory data analysis, including some summary statistics and visualizations, along with some explanation on how they help you learn more about your data. (You can add to these later as you work on your project.)
- The data visualization(s) that you believe will be useful in exploring your question(s). (You can update these later as you work on your project.)


```{r}
# Code goes here
# Code to calculate summary statistics
# Code for a visualization
```
