---
title: "Project memo"
author: "ADA"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(skimr)
```

## Data Clean Up Steps for Overall Data

### Step 1: Cleaning both data sets

```{r clean-demographic}
demographic_clean <- demographic %>%
  # ensure Group is character 
  mutate(Group = as.character(Group)) %>%
  
  # convert only columns that actually contain "%" into numeric
  mutate(across(
    .cols = where(~ any(grepl("%", .x, fixed = TRUE), na.rm = TRUE)),
    .fns  = ~ as.numeric(gsub("%", "", .x))
  )) %>%
  
  # create gender and race columns
  mutate(
    gender = case_when(
      Group %in% c("Men", "Women") ~ Group,
      TRUE ~ NA_character_
    ),
    race = case_when(
      Group %in% c("White", "Black", "Hispanic", "Asian") ~ Group,
      TRUE ~ NA_character_
    )
  ) %>%
  
  # reshape to long format
  pivot_longer(
    cols = -c(Group, gender, race),
    names_to  = "platform",
    values_to = "percent"
  ) %>%
  filter(!is.na(platform)) 

```

```{r clean-social}
social_clean <- social %>%
  mutate(across(where(is.character), 
                ~ as.numeric(ifelse(is.na(.), NA, gsub("%", "", .))))) %>%
  
  # Keep year column, rename if needed
  rename(date = Year) %>%
  
  # Reshape to long format
  pivot_longer(
    cols = -date,
    names_to = "platform",
    values_to = "percent"
  ) %>%
  mutate(year = lubridate::year(lubridate::mdy(date)))
```

### Step 2: Visualization for presentation on 11/13

## Plots

### ggsave example for saving plots

```{r starwars ggsave example}
p1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```



### Plot 1: Platform Usage by Race

#### Data cleanup steps specific to plot 1

These data cleaning sections are optional and depend on if you have some data cleaning steps specific to a particular plot

```{r}

```


#### Final Plot 1

```{r platform-usage-by-race}
ggplot(demographic_clean %>% filter(!is.na(race)), 
       aes(x = platform, y = percent, fill = race)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  labs( title = "Social Media Usage by Platform and Race", 
        x = "Social Media Platform", 
        y = "Usage (%)", fill = "Race" ) + 
  theme_minimal(base_size = 13) + 
  theme( axis.text.x = element_text(angle = 45, hjust = 1),
         plot.title = element_text(face = "bold", size = 16, hjust = 0.5) ) 
ggsave("platform-by-race.png", width = 4, height = 4)

```


### Plot 2: Platform Usage by Gender

```{r platform-usage-by-gender}
ggplot(demographic_clean %>% filter(!is.na(gender)), 
       aes(x = gender, y = percent, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  
  # --- ADD LABELS ON TOP OF BARS ---
  geom_text(
    aes(label = percent),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 2.5
  ) +
  
  scale_fill_manual(values = c(
    "Men" = "#0072B2",
    "Women" = "#CC79A7"
  )) +
  
  # no y-axis labels needed anymore
  scale_y_continuous(
    breaks = c(0, 25, 50, 75, 100),
    limits = c(0, 100),       # give extra headroom for text labels
    expand = expansion(mult = c(0, 0.05))
  ) +
  
  labs(
    title = "Social Media Usage by Platform and Gender",
    x = "Gender",
    y = "Usage (%)"
  ) +
  
  facet_wrap(~ platform, nrow = 3) +
  
  theme_minimal(base_size = 10) +
  theme(
    axis.text.y = element_blank(),   # remove y-axis numbers
    axis.ticks.y = element_blank(),
    panel.border = element_blank(),
    panel.grid.major.y = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 8, angle = 0),
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    legend.position = "none"        
  )

ggsave("platform-by-gender-facet.png", width = 4, height = 4)
```

### Plot 3: ___________

Add more plot sections as needed. Each project should have at least 3 plots, but talk to me if you have fewer than 3.

### Plot 4: ___________ 

